---
title: "Assessment on Statistical analysis for data science(C7081)"
author: "22302900"
date: "2022-11-23"
output: word_document
---

GitHub link: [https://github.com/kksam2705/my-project.git]

# Background

The data set contains several parameters which are considered important during the application for Master's Programs.

The important parameters are GRE score(out of 340), TOEFL score(out of 120), university rating(out of 5), SOP (Statement of Purpose)(out of 5), LOR (Letter of Recommendation)(out of 5), CGPA(out of 10), Research(0 or 1), and a chance of admission(out of 10).

In this data set, 400 entries are included with 8 variables.

# Objective

we are going to Analyze which are all factors that are important for getting Admission to the university.

# Source
This data set was downloaded from Kaggle.
https://www.kaggle.com/datasets/mohansacharya/graduate-admissions


# Methods

# 1. Simple Linear Regression

Simple linear regression is used to estimate the relationship between two quantitative variables. How strong the relationship is between two variables.

# 2. Multiple Regression

Multiple regression is an extension of linear regression into a relationship between more than two variables. In simple linear relation, we have one predictor and one response variable, but in multiple regression, we have more than one predictor variable and one response variable.

# 3. Decision Tree

A Decision tree is a graph to represent choices and their results in form of a tree. The nodes in the graph represent an event or choice and the edges of the graph represent the decision rules or conditions.

# 4. Random Forest

In the random forest approach, a large number of decision trees are created. Every observation is fed into every decision tree. The most common outcome for each observation is used as the final output. A new observation is fed into all the trees and takes a majority vote for each classification model.

# 5. Logistic Regression

Logistic Regression is a regression model in which the response variable (dependent variable) has categorical values such as True/False or 0/1. It actually measures the probability of a binary response as the value of the response variable based on the mathematical equation relating to the predictor variables.


# Setup 

```{r}
setwd("C:/Users/DELL/Downloads/admission into university")
library(readr)
adm_data <- read_csv("adm_data.csv")
names(adm_data)
plot(adm_data)
```

GRE Score, TOEFL Score, and CGPA are important because they are highly correlated.

# Summary of all variables
```{r}
summary(adm_data)
```
Checking the Average score, minimum score, and maximum score of all variables.

## Linear Regression #### 
```{r}
adm.lm <- lm(ChanceofAdmit ~ CGPA, data = adm_data)
adm.lm
```

Make Liner regression with Chance of Admit as an independent variable and CGPA as a dependent variable.

# summary
```{r}
summary(adm.lm)
```


The estimates (Estimate) for the model parameters – the value of the y-intercept (in this case -1.071) and the estimated effect of Chance of Admit on CGPA (0.208).

The final three lines are model diagnostics – the most important thing to note is the p-value (here it is 2.2e-16, or almost zero), which will indicate whether the model fits the data well.

# Plot the data
```{r}
plot(
  ChanceofAdmit ~ CGPA,
  data = adm_data,
  xlab = "Chance of Admit",
  ylab = "CGPA",
  main = "Admission plot",
  col = "blue",
  lwd = 1,
  pch = 20
)

# Add regression line
abline(reg = adm.lm, col = "red", lwd = 3)
```

# Diagnostic plots
```{r}
par(mfrow = c(2, 2))
plot(adm.lm)
```

# Residual distribution
```{r}
par(mfrow = c(1, 1))
hist(residuals(adm.lm), main = "")
```
 
The most important thing to look for is that the red lines representing the mean of the residuals are all basically horizontal and centered around zero. This means there are no outliers.
 
In the Normal Q-Q plot in the top right, we can see that the real residuals from our model form an almost perfectly one-to-one line with the theoretical residuals from a perfect model.
 
Based on these residuals, we can say that our model meets the assumption of homoscedasticity.

# Visualize the results with Graph
```{r}
library(ggplot2)
adm.graph <-
  ggplot(adm_data, aes(x = ChanceofAdmit, y = CGPA)) + geom_point()
adm.graph
```

# Add regression line
```{r}
adm.graph <- adm.graph + geom_smooth(method = "lm", col = "blue")
adm.graph
```

# Conclusion
From these results, we can say that there is a significant positive relationship between the Chance of Admit and CGPA (p-value < 0.001), with a 0.208-unit (+/- 0.01) increase in Chance of Admit for every unit increase in CGPA.

Since there is a significant positive relationship between the Chance of Admit and CGPA so, we can reject the null hypothesis.

## Multiple regression ####
 Check for linearity.
 
We can check this using two scatter plots: one for GRE Score and Chance of admission, and one for TOEFL Score and Chance of admission.

 ```{r}
plot(adm_data$ChanceofAdmit, adm_data$GREScore)
plot(adm_data$ChanceofAdmit, adm_data$TOEFLScore)
adm.multi <-
  lm(ChanceofAdmit ~ GREScore + TOEFLScore, data = adm_data)
adm.multi
```

# summary

```{r}
summary(adm.multi)
```

The estimated value of chance of admission on the GRE Score is 0.0058, while the estimated value of the TOEFL Score is 0.0094.

# Check for homoscedasticity

```{r}
par(mfrow = c(2, 2))
plot(adm.multi)
par(mfrow = c(1, 1))
```
# Create the Equation for a Regression model

# y = a + b1x1 + b2x2 +...bnxn

# Apply Equation for predicting New Values

we can predict the chance of Admission when we set new values to GRE Score and TOEFL Score.


```{r}
Y = -2.128+(0.005814)*325+(0.009412)*105
Y

```

# Conclusion

If the person with GRE Score is 325 and TOEFL Score is 105 and his chance of Admit is 0.74%.

Since there is a positive relationship between the GRE Score, TOEFL Score, and the Chance of admitting so, we can reject the null hypothesis.


# Visualize the results with graph

```{r}
plotting.data <- expand.grid(
  GREScore = seq(min(adm_data$GREScore), max(adm_data$GREScore), length.out =
                   30),
  TOEFLScore = c(
    min(adm_data$TOEFLScore),
    mean(adm_data$TOEFLScore),
    max(adm_data$TOEFLScore)
  )
)
plotting.data$predicted.y <-
  predict.lm(adm.multi, newdata = plotting.data)

plotting.data$GREScore <- round(plotting.data$GREScore, digits = 2)

multi.plot <-
  ggplot(adm_data, aes(x = GREScore, y = ChanceofAdmit)) + geom_point()
multi.plot

multi.plot <- multi.plot +
  geom_line(data = plotting.data,
            aes(x = GREScore, y = predicted.y, color = TOEFLScore),
            size = 1.25)
multi.plot

multi.plot <- multi.plot + theme_bw() +
  labs(
    title = "Chance of Admit \n as a function of GREScore to work and TOEFLScore",
    x = "GRE Score",
    y = "Chance of Admit",
    color = "TOEFL Score" )
multi.plot
```

## Decision tree ####

```{r}
library(randomForest)
library(party)
input.data <- adm_data[c(1:105),]
output.tree <-
  ctree(ChanceofAdmit ~ CGPA + GREScore + TOEFLScore, data = input.data)
plot(output.tree)
```


# conclusion

From the decision tree, we can conclude that anyone whose CGPA is more than 8.8 and whose GRE Score is more than 311 will have a high chance of Admission.

## Random forest ####

```{r}
library(randomForest)
library(party)
adm.reg <- randomForest(ChanceofAdmit ~ ., data = adm_data)
print(adm.reg)
round(importance(adm.reg), 2)
```

# Conclusion

From the random forest, we can conclude that the CGPA, GRE Score and TOEFL Score are the important factors for getting admission into the university.

## Logistic Regression ####

```{r}
glm.adm <- glm(ChanceofAdmit ~ ., data = adm_data)
glm.adm
summary(glm.adm)
```

# Coefficients

Intercept coefficient(-1.2594325)

GRE Score coefficient(0.0017374)

TOEFL Score Coefficient(0.0029196)

LOR Coefficient( 0.0223531)

CGPA Coefficient(0.1189395)

Research Coefficient(0.0245251)

These are all the factors that impact the Chance of Admit.


# p-value

The p-value of the University Rating and SOP are so high (more than 0.05) which implies that insignificant to the Chance of Admit.
 
while other factors such as GRE Score, TOEFL Score, CGPA, and LOR with low p-value(less than 0.05) are significant to the Chance of Admit.

# Predict()

```{r}
glm.predict <- (predict(glm.adm, type = "response"))
glm.predict
```

## Result ####

By performing various statistical analyses CGPA, GRE Score, and TOEFL Score, are all crucial factors that obtain the Chance of Admit to the University.

## Literature citation ####

Mohan S. Acharya, Asfia Armaan, Aneeta S Antony: A Comparison of Regression Models for Prediction of Graduate Admissions, IEEE International Conference on Computational Intelligence in Data Science 2019.


```{r example-code-chunk, include = TRUE}